### 2016-05-16 Paper/paper.tex
full draft up to methods

\begin{abstract}
  This report reviews the spectral analysis method for modeling ocean surface drifters proposed by \citet{Sykulski2016}. Previous methods to model drifters are reviewed along with the authors' model.   Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are considered.  An evaluation of the strengths and weaknesses of this method concludes the review. 
  \end{abstract}

@@

Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation and is also sometimes referred to as the complex-valued stationary autoregressive process. This stochastic differential equation creates a process  that is both stationary and Markovian. The parameter $f_{0}$ represents the frequency of inertial oscillations. The dampening parameter $c > 0$  guarantees that the process is mean-reverting, meaning that it will always eventually return to its mean position, $z = 0$.   The final term of the differential equation is a positive constant $A$ multiplied by $dQ(t)$.  The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes, is a standard complex Wiener process. We remind the reader that Wiener processes are stochastic models originally used to represent the random motion of a particle. Such a process $\{Q(t, \omega): t \geq 0\}$ on the space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $|\tau|$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

@@

\subsubsection{Turbulent Background}
The drifters' time series are also marked by large-scale turbulence (read and cite Rhines 1979). Many potential models have been considered for this background....Lit review here ... However, these models have been \textbf{\it{integer order}}, meaning the velocity, acceleration, and hyperacceleration must be Markovian, a constraint previous studies (READ AND CITE) indicate is not appropriate in all cases. Thus, the authors propose using the Mat\'{e}rn covariance structure \citep{Gneiting2012}, a model which has gained prominence in the spatial statistics literature due to its flexibility. In this context, the Mat\'{e}rn model proves to be useful, since it allows for markovianity, but does not require it.  The Mat\'{e}rn  autocovariance and power spectral density are
\begin{align}
\label{eq:maternAC}
s^{(m)}(\tau) &= \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\\
\label{eq:maternPSD}
S^{(m)}(\omega) &= \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}
\end{align}
where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}. The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.

>>>>

\begin{abstract}
  This report reviews the spectral analysis method for modeling ocean surface drifters proposed by \citet{Sykulski2016}. Previous methods to model drifters are discussed along with the authors' model.   Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are reviewed.  To conclude the review, we evaluate the strengths and weaknesses of this method. 
  \end{abstract}

@@

Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation and is also sometimes referred to as the complex-valued stationary autoregressive process. This stochastic differential equation creates a process  that is both stationary and Markovian. The parameter $f_{0}$ represents the frequency of inertial oscillations. The dampening parameter $c > 0$  guarantees that the process is mean-reverting, meaning that it will always eventually return to its mean position, $z = 0$.   The final term of the differential equation is a positive constant $A$ multiplied by $dQ(t)$.  The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$ is a standard complex Wiener process, defined to be such that $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes, . We remind the reader that Wiener processes are stochastic models originally used to represent the random motion of a particle. Such a process $\{Q(t, \omega): t \geq 0\}$ on the space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $|\tau|$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

@@

\subsubsection{Turbulent Background}
The drifters' time series are also marked by the ocean's large-scale turbulence. The underlying physics describing this red noise process is summarized in \citet{Rhines1979}. Many models for ocean turbulence have been proposed, including several stochastic models \citep{Lacasce2008}. Within the oceanography literature, the stochastic models are designated  first-order \citep{Griffa1995, Falco2000}, second-order \citep{Sawford1991}, and third-order corresponding to  whether the velocity, acceleration, or hyperacceleration is modeled with a Markov process. However, calculation of the spectral slopes in observed data indicate that this class of models, referred to as \textbf{\it{integer order}},  does not always adequately reflect the ocean's behavior \citep{Rupolo1996, Sanderson1991}. Rather, \textbf{\it{fractional power models}}, such as fractional Brownian motion (FBM), would better reflect the observed behavior in some cases.

Towards this end, the authors propose using a Mat\'{e}rn covariance structure \citep{Gneiting2012} for the turbulent background. The Mat\'{e}rn model has gained prominence in the spatial statistics literature due to its flexibility. In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers. The Mat\'{e}rn model is defined either by its autocovariance or power spectral density
\begin{align}
\label{eq:maternAC}
s^{(m)}(\tau) &= \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\\
\label{eq:maternPSD}
S^{(m)}(\omega) &= \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}
\end{align}
where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}. The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.

The authors emphasize that the turbulent background models previously proposed are special cases of the Mat\'{e}rn model. This means that these models can still be obtained under the current model if model fitting leads to particular parameter estimate values. They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean. Figure ~\ref{fig: fbmMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.

\begin{figure}[hb]
	\label{fig: fmbMat}
  \centering
    \includegraphics[width=.5\textwidth]{fig3.pdf}
        \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes. Top Left: The eastward component the velocity; Top Right: The eastward and northward components of the vecocity; Bottom: Periodograms on the decibel scale}
\end{figure}
